---
title: "Mapping species distribution"
author: "Mauro Lepore"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 6
vignette: >
  %\VignetteIndexEntry{Mapping species distribution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# hadley's settings
set.seed(1014)
options(digits = 3)

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)

library(ggplot2)
update_geom_defaults("text", list(size = 14.5))
```

# Overview

This vignette shows you how to map species distribution with `map_sp()` and `map_sp_pdf()` from the __forestr__ package.

```{r}
# To install from a private repo, generate a personal access token (PAT) in
# https://github.com/settings/tokens and supply to this argument.
GITHUB_PAT <- "your token"
# install_github("forestgeo/forestr", auth_token = GITHUB_PAT)
library(forestr)

# Also using the package dplyr for easier data manipulation
# install.packages(dplyr)
library(dplyr)
# Print only a few rows of tibbles (modern dataframes) to save space
options(dplyr.print_min = 6, dplyr.print_max = 6)
```

Loading data for examples.

```{r}
# Converting dataframe to tibble for better printing
example_data <- as_tibble(forestr::bci12t7mini)

# Let's find the top-3 most abundant species
sp_n <- count(example_data, sp)
sp_arranged <- arrange(sp_n, desc(n))
sp_top3 <- sp_arranged$sp[1:3]
sp_top3

# For a small example, keeping data of only the 3 most abundant species
census <- filter(example_data, sp  %in% sp_top3)
census
```

The next section shows you how to explore data quickly. You will produce good quality maps with available defaults. The section after the next one shows you how to customize you maps by making some common and less common changes. And the last section shows you how to fine tune your maps.

# Exploring Data Quickly: Using Defaults

## Mapping Species Distribution

Although the output of `map_sp()` and `map_sp_pdf()` is printed to different devices -- i.e, the screen or a .pdf file --  those functions are almost identical. They are documented jointly and share the same arguments (except for the `file` argument that only belongs to `map_sp_pdf()`. (These functions are not merge into a single function because their outputs are different; functions that produce different output tend to cause more problems that those which output is consistently the same).

```
?map_sp
# Or
?map_sp_pdf
```

Notice that only the first two arguments of these functions are strictly necessary.

```{r}
# Print to screen
map_sp(census, c("hybapr", "faraoc"))
```

```{r}
# Defaults to save in working directory. Returns invisibly.
map_sp_pdf(census, c("hybapr", "faraoc"))
```

```{r fig.align="default", out.width="45%", fig.widh=(6 * 0.45 / 0.7)}
# Force printing the invisible output
print(map_sp_pdf(census, c("hybapr", "faraoc")))

# (Printing side by side to save space)
```

Keep in mind that your census data must have specific variable-names:

```{r, error=TRUE}
# Show behaviour with wrong name
with_wrong_name <- rename(census, SP = sp)

# This will fail because the data set has name `SP`, not `sp`
map_sp(with_wrong_name, c("hybapr", "faraoc"))
```

## Adding Elevation

You may have elevation data from your site.

```{r}
my_elev <- forestr::bci_elevation
my_elev
```

Note that the elevation data set must have very specific variable names; or it will fail:

```{r, error=TRUE}
# This will fail because the names of `my_elev` are not correct
map_sp(census, c("hybapr", "faraoc"), elevation = my_elev)
```

If necessary, rename your variables:

```{r}
# Renaming data and trying again
elev <- rename(my_elev, gx = x, gy = y)
map_sp(census, c("hybapr", "faraoc"), elevation = elev)
```

# Dealing with Overplotting

Sometimes a species is so abundant that its distribution overplots.

```{r}
# Fake over abundant species
crowded <- dplyr::tibble(
  sp = sample(c("species1"), 10000, replace = TRUE),
  gx = sample.int(1000, 10000, replace = TRUE),
  gy = sample.int(500, 10000, replace = TRUE)
)
map_sp(crowded, c("species1"))
```

To reduce overplotting, you can do a number of things. First, you can try reducing the size or opacity of the points, or changing their shape.

```{r fig.align="default", out.width="45%", fig.widh=(6 * 0.45 / 0.7)}
# smaller and semi transparent
map_sp(crowded, c("species1"), size = 1, alpha = 2/10)

# Smaller and open
map_sp(crowded, c("species1"), size = 1, shape = 21)

# (Printing side by side to save space)
```

If your research question allows, you can try subsampling your data.

```{r}
# A random sample of 2000 stems
smaller <- sample_n(crowded, 2000)
map_sp(smaller, c("species1"), shape = 21)
```

# Customizing Your Maps

You may want to customize your maps. For example, you may want to change some properties of the points or elevation lines (e.g. colour); or the general theme of the map. These are relatively common changes.

## Most Common Changes

### Changing Points' Properties

Both `map_sp()` and `map_sp_pdf()` have the argument `...`, which lets you pass arguments to `ggplot2::geom_point()` to change points' properties such as the shape, size, fill, colour, and stroke. Read `?ggplot2::geom_point()` to learn all your options; here you can see some common ones:

```{r}
map_sp(census, "hybapr",
  # Arguments passed to ggplot2::geom_point()
  size = 4, shape = 22, fill = "green", colour = "black", stroke = 2)
```

### Changing Properties of the Elevation Lines

You can also customize the elevation lines. This example shows all the available options:

* Change line size;
* set colours to represent low and high elevation;
* set the number of intervals in which the elevation variable is to be cut (note that the cut results in ugly numbers, with many decimals; also, with increasing bins' number the function takes longer to run).

```{r}
map_sp(census, "hybapr", 
  elevation = elev, line_size = 1, low = "red", high = "blue", bins = 4)
```


### Changing the Theme

You can change the general look of your map by changing its theme. A number of pre-set themes are available in the __ggplot2__ package (see `?ggplot2::theme_bw()`).

```{r, fig.align="default", out.width="45%", fig.widh=(6 * 0.45 / 0.7)}
map_sp(census, "hybapr", theme = ggplot2::theme_classic())
map_sp(census, "hybapr", theme = ggplot2::theme_dark())

# (Printing side by side to save space)
```

## Less Common Changes

Here are some changes that you may need less often but are still good to keep in mind.

### Change File Name

When printing to .pdf, you can provide a custom file name (including the path to a directory); but note the file name must end in _.pdf_ or it will be replaced by a default name:

```{r, error=TRUE}
spp <- c("hybapr", "faraoc")

map_sp_pdf(census, spp, elevation = elev, file = "bad-name1")

map_sp_pdf(census, spp, elevation = elev, file = "bad-name2.png")

```

This is OK. (Note that you can suppress messages)

```{r}
map_sp_pdf(census, spp, elevation = elev, file = "good-name.pdf")

# Same without any message
suppressMessages(
  map_sp_pdf(census, spp, elevation = elev, file = "good-name.pdf")
)
```



### Change x and y Limits

Occasionally you may need to change the map limits. The map limits default to the range 0-max, where max is the maximum value of the variable `gx` or `gy` in the census data. In the unlikely case where no stem occurs at the maximum limits of your plot, the default would result in a map that spans less than the defined limits of the site's plot.

```{r, fig.align="default", out.width="45%", fig.widh=(6 * 0.45 / 0.7)}
# Fake case: The maximum-limits of the site's plot are gx = 1000 and gy = 500;
# but all stems in the census occurr at lower gx and gy limits
odd_data <- filter(census, gx < 800, gy < 200)
two_species <- unique(odd_data$sp)[1:2]
map_sp(odd_data, two_species)

# (Printing side by side to save space)
```

You can overwrite the default limits with the arguments `xlim` and `ylim`:

```{r, fig.align="default", out.width="45%", fig.widh=(6 * 0.45 / 0.7)}
map_sp(odd_data, two_species, xlim = c(0, 1000), ylim = c(0, 500))

# (Printing side by side to save space)
```

# Fine Tunning

This last section shows you how to fine tune your plots. For this, you will need a couple of functions from packages other than __forestr__.

## Multiple Plots in One Page

To plot multiple maps on one page you can use `gridExtra::maparrange()`:

```{r}
library(gridExtra)

all_species <- unique(census$sp)
maps <- map_sp(census, all_species)
multipaged <- marrangeGrob(maps, nrow = 1, ncol = 2)
multipaged
```

The multi-paged maps can also be saved to a .pdf.

```{r}
# Saving to .pdf: Option 1
ggplot2::ggsave("multi-paged.pdf", multipaged)

# Saving to .pdf: Option 2
pdf()
multipaged
dev.off()
```

It is also possible to save the multi-paged maps to devices other than .pdf -- the png() device, for example, will automatically create one file per page.

```{r}
# Saving to .png; this will create multiple files.
png()
multipaged
dev.off()
```

## Customizing the Theme

Before, you used a pre-made theme. But you can create your own theme with `ggplot2::theme()`, save it as an object, and pass that object to `map_sp()` or `map_sp_pdf()`. 

```{r}
# Showing only a few options; see all the options with ?ggplot2::theme()
my_theme <- ggplot2::theme(
  text = element_text(size = 25, face = "bold.italic", colour = "white"),
  plot.background = element_rect(fill = "black"),
  plot.margin = margin(2, 2, 2, 2, "cm"),
  strip.background = element_rect(fill = "darkgreen"),
  strip.text = element_text(colour = "white"),
  # make grid to dissapear by matching background colour
  panel.background = element_rect(fill = "lightgreen"),
  panel.grid.minor = element_line(colour = "lightgreen"),
  panel.grid.major = element_line(colour = "lightgreen")
)
map_sp(census, "hybapr", theme = my_theme)
```

## Extending with ggplot2

Finally, you can also use the output of `map_sp()` and `map_pdf()` as the starting point for further customization with __ggplot2__ (the next section provides resources to learn __ggplot2__). That may be convenient for minor extensions; for larger ones you may prefer to start from scratch. Note that `map_sp()` and `map_sp_pdf()` output not just one plot but a list of plots. That is important to keep in mind if you want to iterate over all the plot-elements of the output list. For example, this is how you would add a new layer to one element of the plots' list:

```{r}
p0 <- map_sp(census, c("hybapr", "faraoc"))

# Adding a new layer to one element of the plots' list
p0[["hybapr"]] + geom_vline(aes(xintercept = 300), colour = "red")
```

And this is how you can apply new layers to all the plot-elements of a list (note that the function `+` is applied to each item of the plots' list):

```{r fig.align="default", out.width="45%", fig.widh=(6 * 0.45 / 0.7)}
p1 <- lapply(p0, `+`, geom_vline(aes(xintercept = 300), colour = "red"))
p1

p2 <- lapply(p1, `+`, geom_hline(aes(yintercept = 400), colour = "blue"))
p2

# (Printing side by side to save space)
```

## Learning More

There are many visualization systems in R. The system built in `map_sp()` and `map_sp_pdf()` is that of the __ggplot2__ package. __ggplot2__ is not an arbitrary collection of functions; it is one implementation of [The Grammar of Graphics](https://goo.gl/E514ar). Good places to learn __ggplot2__ include [this website](https://goo.gl/tVvLQZ); this specialized, paid book: [ggplot2: Elegant Graphics for Data Analysis](https://goo.gl/R6wJk9); and this general, free book [R for Data Science](http://r4ds.had.co.nz/).
