---
title: "Plot Species Distribution"
subtitle: "An example for maximum flexibility"
author: "Mauro Lepore"
date: "2017-08-03"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 6
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

# > If you want to make sure the font size is consistent across all your 
# figures, whenever you set out.width, you'll also need to adjust fig.width to 
# maintain the same ratio with your default out.width. For example, if your 
# default fig.width is 6 and out.width is 0.7, when you set out.width = "50%" 
# you'll need to set fig.width to 4.3 (6 * 0.5 / 0.7).
# --http://r4ds.had.co.nz/

set.seed(1014)
options(digits = 3)

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  out.width = "90%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)

options(dplyr.print_min = 6, dplyr.print_max = 6)
```

```{r pkg}
# Useful packages
library(ggplot2)
library(dplyr)
```

```{r cns_data_sub}
# Data
library(bci)
# Data for a minimal example.
cns_data <- bci::bci12full7
# Filtering 3 species only
species_selected <- unique(cns_data$sp)[1:3]
cns_data_sub <- filter(cns_data, sp %in% species_selected)
# Sampling only some rows
cns_data_sub <- sample_n(cns_data_sub, 10000)
```

# Most flexible approach

Set plot limits for all plots to have consistent size.

```{r plot-limits}
xlimits <- c(0, max(cns_data$gx, na.rm = TRUE))
ylimits <- c(0, max(cns_data$gy, na.rm = TRUE))
```

Make the plot you want for each species; then wrap it with a function. __This step is what makes this approach flexible; you can do whatever plot you want.__

```{r plot-sp}
plot_sp <- function(cns_data_sub) {
  ggplot(data = cns_data_sub, aes(x = gx, y = gy)) +
    geom_point() +
    facet_grid(. ~ sp) +
    coord_fixed(xlim = xlimits, ylim = ylimits) +
    theme_bw()
}
```

Make a list of data frames, split by species; then apply your plotting function to each element of the data frames list.

```{r plot-list}
split_by_sp <- split(cns_data_sub, cns_data_sub$sp)
plot_list <- lapply(split_by_sp, plot_sp)
```

Save the resulting list of plots to your favorite device.

```R
pdf()
plot_list
dev.off()
```

For example, the code above produced the plots shown below, but on a single .pdf file with one species plotted on a different page.

```{r show-saved, fig.align="default", out.width = "30%", fig.widh = (6 * 0.3 / 0.9)}
plot_list
```

# Quickest approach

When you can sacrifice flexibility for speed, you may try `lapply_plot_sp()`:

```{r show-quick-default, fig.align="default", out.width="30%", fig.widh=(6 * 0.3 / 0.9)}
# The package try is where I try new things that I may or may not develop.
library(try)

# Selecting all species in the example dataset
all_species <- unique(cns_data_sub$sp)
lapply_plot_sp(species = all_species, cns_data = cns_data_sub)
```

With `lapply_plot_sp()`, you can (only) vary the looks of the plotted points:

```{r show-quick-tweaked, fig.align="default", out.width="30%", fig.widh=(6 * 0.3 / 0.9)}
lapply_plot_sp(
  species = all_species, cns_data = cns_data_sub,
  # Change the look of the points  passed to ?geom_point
  color = "blue", size = 3, shape = 1
)
```












