---
title: "Plot Repulsive Tags from Sinharaha"
author: "Mauro Lepore"
date: "2017-09-26"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 6
vignette: >
  %\VignetteIndexEntry{Plot repulsive tags from Sinharaha}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# hadley's settings
set.seed(1014)
options(digits = 3)

knitr::opts_chunk$set(
  echo = TRUE,  # {mine}
  comment = "#>",
  collapse = TRUE,
  # cache = TRUE,
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)

options(dplyr.print_min = 6, dplyr.print_max = 6)
```

# First attempt

This code was originally used to output the maps that were first printed.

```
# The sinharaja package is local and the data set sinh_q20 is private.
list_of_dataframes <- sinharaja::sinh_q20

# Filder selected quadrats
path <- "../sinharaja/data-raw/from_shameema/"
file_nm <- "Quadrat abundances 2017.to_check.xlsx"
selected_quads <- readxl::read_xlsx(paste0(path, file_nm)) %>%
  select(Quad_Name, QuadCheck) %>%
  filter(QuadCheck %in% c(1.5, 2)) %>%
  unique() %>%
  pull(Quad_Name)
selected_list_of_dataframes <- list_of_dataframes[selected_quads]
# Confirm
identical(selected_quads, names(selected_list_of_dataframes))

# Prepare and pdf
prepared <- prep_repulsive_tags(selected_list_of_dataframes)
plot_list <- lapply_plot_repulsive_tags(prepared, site_name = "Sinharaja 2017")

pdf(onefile = TRUE, paper = "a4", width = 11, height = 11)
plot_list
dev.off()
```

# Setup 

```{r}
library(tidyverse)
library(try)
```



# Second attempt

Here I use subquadrat information 

From ViewFullTable, here I get information for each tree including its subquadrat, and position. ViewFullTable  also has information on tree status, which I could use directly to plot the latest status of each tree. This seems to be exactly what a general function should do. For that, the code below should be adapted to use the ViewFullTable specifically.

This vignette, however, shows a special case. Here, the tree status that I am plotting comes not from ViewFullTable but from data that Shameema provided. Those data from Shameema contain a particular status that Stuart wants to recheck. That status Shameema represented with some symbols. I am unsure how Shameema calculated that status because it does not match the ViewFullTable directly. This  is  something I asked Shameema (via https://goo.gl/NzoxqY), but as per Nov 4th, 2017, she didn't reply.

## Merge ViewFullTable with Shameema's list of dataframes

* ViewFullTable has the correct information of the subquadrat where each tree should be.
* Shameema's list of dataframes has the data on symbol and status

Import data.

```{r}
vftbl <- sinharaja::sinh_vftbl_selected
vftbl
```

Raname data to match the needs of existing functions.

```{r}
n_original <- nrow(vftbl)

renamed <- vftbl %>% 
  select(-id, -StemTag, -Status) %>% 
  rename(
    tag = Tag,
    subquadrat_vftbl = Subquadrat,
    quadrat_vftbl = QuadratName,
    lx = QX,
    ly = QY
  ) %>% 
  unique()
renamed
```

Get symbols from Shameema's list of dataframes

```{r}
sham_df <- sinharaja::sinh_q20 %>% 
  reduce(full_join) %>% 
  as_tibble()
# allow merging by tag
sham_df <- sham_df %>% 
  mutate(
    status = as.character(status),
    tag = str_pad(tag, width = 6, side = "left", pad = "0")
  ) %>% 
  select(-lx, -ly, -subtag) %>% 
  unique()
sham_df
```

Merge selected data (quadrats chosen by Stuart) from ViewFullTable with status and symbols by Shameema

```{r}
sinh_df <- left_join(renamed, sham_df)
sinh_df
```

Split the data

```{r}
splitted <- sinh_df %>% split(.$quadrat_vftbl)
```

Table problematic trees

```{r}
splitted %>% 
  map(as.data.frame) %>% 
  map(filter, lx >= 20 | ly >= 20) %>% 
  map(arrange, tag) %>% 
  map(select, tag, subquadrat_vftbl)
```



## Prepare and plot

```{r}
# Prepare and pdf

# funs ----
paginate <- function(x) {
  dplyr::mutate(x, subquadrat =
      case_when(
        subquadrat_vftbl == 11 ~ 1,
        subquadrat_vftbl == 12 ~ 1,
        subquadrat_vftbl == 21 ~ 1,
        subquadrat_vftbl == 22 ~ 1,

        subquadrat_vftbl == 31 ~ 2,
        subquadrat_vftbl == 32 ~ 2,
        subquadrat_vftbl == 41 ~ 2,
        subquadrat_vftbl == 42 ~ 2,

        subquadrat_vftbl == 34 ~ 3,
        subquadrat_vftbl == 33 ~ 3,
        subquadrat_vftbl == 44 ~ 3,
        subquadrat_vftbl == 43 ~ 3,

        subquadrat_vftbl == 14 ~ 4,
        subquadrat_vftbl == 13 ~ 4,
        subquadrat_vftbl == 24 ~ 4,
        subquadrat_vftbl == 23 ~ 4,
      )
    )
}

add_sqds <- function(df) {
  paginate(df) %>% 
  dplyr::group_by(subquadrat) %>% 
  dplyr::mutate(sqds = paste0(unique(sort(subquadrat_vftbl)), collapse = "-")) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(sqds, everything())
}

prep_repulsive_tags <- function(df_list) {
  df_list %>% 
    purrr::map(add_sqds) %>%
    purrr::map(add_latest_tree_status) %>%
    purrr::map(paginate) %>%
    purrr::map(dplyr::rename, quadrat = quadrat_vftbl) %>%
    purrr::map(add_subquad_limits) %>%
    purrr::map(dplyr::mutate,
      id = paste0("Q. ", quadrat, " SQ. ", sqds, " (p. ", subquadrat, ")")
    ) %>%
    purrr::map(dplyr::select, id, subquadrat, dplyr::everything()) %>%
    purrr::map(dplyr::select,
      id, tag, lx, ly, latest_tree_status, x1, x2, y1, y2, everything()
    ) %>% 
    reduce(full_join) %>% 
    # Add status to tag because some points dissapear from plot but tags  
    # persist
    mutate(
      tag = case_when(
        latest_tree_status == "Alive" ~ paste0(tag, "_"),
        latest_tree_status == "Dead" ~ paste0(tag, ".")
      )
    ) %>% 
    split(., .$id)
}



# action ----

prep_list <- splitted %>% prep_repulsive_tags()
plot_list <- prep_list %>% 
  lapply_plot_repulsive_tags(site_name = "Sinharaja 2017")

pdf(onefile = TRUE, paper = "a4", width = 11, height = 11)
plot_list
dev.off()
```



Just two extreeme quadrats.

```{r}
df <- splitted %>% reduce(full_join)
df %>% filter(lx == max(lx) | ly == max(ly))

prep_df <- prepared %>% reduce(full_join)
prep_df <- prep_df %>% 
  filter(quadrat %in% c("1115", "2105"))


prep_list <- split(prep_df, prep_df$id)
plot_list <- lapply_plot_repulsive_tags(prep_list, site_name = "Sinharaja 2017")

pdf(onefile = TRUE, paper = "a4", width = 11, height = 11)
plot_list
dev.off()


```





