---
title: "Avoid text overlap"
subtitle: "Example using data from Sinharaja"
author: "maurolepore@gmail.com"
date: "2017-09-20"
output:
  rmarkdown::html_document:
    toc: true
    toc_depth: 6
---

```{r setup, include=FALSE}
# > There is one chunk name that imbues special behaviour: setup. When you're in
# a notebook mode, the chunk named setup will be run automatically once, before
# any other code is run.
# --http://r4ds.had.co.nz/r-markdown.html#chunk-options



# Load packages (including data-packages) ----

library(tidyverse)
library(ggrepel)
library(try)



# Load un-packaged dependencies ----





# Figures ----

# > If you want to make sure the font size is consistent across all your
# figures, whenever you set out.width, you'll also need to adjust fig.width to
# maintain the same ratio with your default out.width. For example, if your
# default fig.width is 6 and out.width is 0.7, when you set out.width = "50%"
# you'll need to set fig.width to 4.3 (6 * 0.5 / 0.7).
# --http://r4ds.had.co.nz/

set.seed(1014)
options(digits = 3)

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  out.width = "90%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = 1,  # 1 / phi
  fig.show = "hold"
)

options(dplyr.print_min = 6, dplyr.print_max = 6)

# customize ggplot

update_geom_defaults("point",
  list(shape = 21, colour = "black", size = 2, stroke = 1)
)
update_geom_defaults("text", list(size = 2))
```

# Goal

To clearly plot information that help workers to identify specific trees on the field. Previous drafts were too crowded, with labels that overlapped and impaired reading the information clearly.

For each quadrat we want to output 4 plots, each showing 1/4 of the quadrat's area. Each of those sub-quadrat we want to plot in one separate page.

# Approach

* Overview the entire data to understand it.
* Pull one data frame.
    * Overview it.
    * Draft a plot quickly to see it we are heading in the right direction.
    * Divide it in 4 subplots.
    * Plot avoiding labels overploting.
        * aesthetics: 
            * `x`, `y` map to position in the quadrat, i.e. `lx`, `ly`.
            * text maps to `tag` id
            * shape or colour map to symbol



# Explore entire data set

> (To Stuart) Since there are 67 quadrats in your check list there will be 268 pages (4 per quad)
--Shameema

The name of each data frame, and their number (67) confirm they correspond to the name (or ID) of each quadrat.

```{r}
names(sin_q20)
```

Let's confirm that all 67 data frames in the list have the same variable names.

```{r}
# Yes. All data frames have exactly the same name.
unique(lapply(sin_q20, names))
```

Now let's overview the data set, paying special attention to the type of each variable.

```{r}
# Show only part of the entire data set to save space
n_cols_per_df <- unlist(unique(lapply(sin_q20, ncol)))
str(sin_q20[1:3], list.len = n_cols_per_df)
```

`status` should probabbly be a character variable (data dictionary https://goo.gl/i6kHiQ) but here it is a factor.

```{r}
# map() is like lapply with some handy shortcuts
map(sin_q20, "status") %>% 
  map(unique) %>% 
  head()
```

The reason why `status` is a factor is probably because the data frames where built using the default `stringsAsFactors = TRUE`. Because it may cause problems, let's fix this now.

```{r}
convert_status_to_character <- function(x) {
  x$status <- as.character(x$status)
  x
}
sin_q20 <- lapply(sin_q20, convert_status_to_character)

# Showing only a few data frames to save space
str(sin_q20[1:3], list.len = n_cols_per_df)
```



# Explore each data set

From the entire data set, pull out only one data frame and overview it

```{r}
first_df <- as.tibble(sin_q20[[1]])
first_df
```



# Plot

Lets filter the lower left quarter of one quadrat and plot it. Later I can apply the same approach to all other quarters of each quadrat, and to all other quadrats. 

```{r}
# Size of points, text and labels
point_size <- 1
id_size <- 2
theme <- theme(
  panel.grid.major = element_line(colour = "black"),
  panel.grid.minor = element_line(linetype = "dotted", colour = "black"),
  panel.background = element_rect(fill = 'white')
  )






# lower left subsection
first_ll <- filter(first_df, lx <= 10, ly <= 10)

ggplot(first_ll, aes(x = lx, y = ly, shape = as.factor(symbol))) +
  geom_point(size = point_size) +
  geom_text_repel(aes(label = tag), size = id_size) +
  theme(legend.position = "top") +
  theme
```

__Please let me know how to make this more fit to your needs. once happy, I'll apply the same style to all plots.__

__Mauro__

